# Information to set you up for your semester {-}

## Course outline {-}

Living systems are the most complex systems in science, and biology is naturally variable and noisy due to its many internal and external influences. For these reasons, it is difficult to make inferences from and predictions about biological systems. Understanding biology requires computational skills to effectively analyse and interpret data, and multidisciplinary research approaches are becoming more common as a critical key to solving many of the complex problems of studying life and living organisms in today’s world. So, contrary to the popular undergrad biology student beliefs, statistics, mathematics and computational skills are essential in a biologist’s toolkit.

To understand modern biological research and findings, and to participate in this research (and get jobs!), skills in working with and visualising data, learning from data using models, and generating data using simulations of models are crucial. These might be classic statistical models, mathematical models, or inference with process-based models. Biologists also need to be careful and critical thinkers about data and how it is acquired, as well as think critically about the models that we use to try to simplify, and thereby understand, the incredible complexity of biology.

BIOSCI 220 Quantitative biology will introduce you to the programming language R to develop the aforementioned skills, with no coding experience assumed or expected. The aim is to give beginners the confidence to continue learning R and not be afraid of statistics and mathematics! 


## Your teaching team, Semester 1, 2024 {-}

### Module 1: Data exploration and statistical inference {-}
**Weeks 1 - 3** Dr Charlotte Jones-Todd
**Weeks 4 - 6** Jenn Jury (course coordinator)

### Module 2: Mathematical modelling in biology {-}
**Weeks 7 - 9** Dr Nobuto Takeuchi

### Module 3: Model-based inference and critical thinking about models {-}
**Weeks 10 - 12** Assoc. Prof. Emma Carroll


## Assessment structure {-}

```{r include=FALSE}
assessment <- matrix(c("2, 3, 5, 6, 8, 9, 11, 12", 5, 40, "2 - 12 inclusive", 1, 10, 7, 20, 20, "Exam period", 30, 30),ncol=3,byrow=TRUE)
colnames(assessment) <- c("Week of semester","Weighting per assessment (%)","Total weighting (%)")
rownames(assessment) <- c("Laboratory","Weekly quiz","Test", "Exam")
assessment <- as.table(assessment)
```
```{r echo=FALSE}
knitr::kable(assessment, align = "cccc", "simple")
```

Note: The test will be held on **Tuesday 12th September 6.30 - 8.15pm NZST**. **The test will cover all of Module 1 taught material only. The exam will cover both Module 2 & 3 taught material only.** The delivery mode for the test and the exam is an on-campus, invigilated Inspera assessment.


## Lectures {-}

Lectures are every week on Mondays and Wednesdays 8-9 am in 260-098 (Owen G Glenn building, 098). We strongly encourage you to attend the lectures in-person as there will be in-class activities to assist you with your learning.\
Lecture recordings will be available in CANVAS via the Panopto video tab as soon as possible after the lecture has been delivered.


## Laboratories (5% each, 40% towards final grade) {-}

Labs will be held in 106-014 (Biology Building, Room 014) during your scheduled SSO lab time, which will be one of Monday 2-5pm, Tuesday 10am-1pm, Tuesday 2-5pm, Thursday 2-5pm or Friday 10am-1pm.\
There are eight (8) labs in total. Labs are held in weeks 2, 3, 5, 6, 8, 9, 11 and 12. (i.e. in each three week lecture block there are labs in the last two weeks).
Labs are three (3) hours and the material is designed to be completed in this time. 
Each lab is worth 5%, totalling 40% towards your final grade.
You are scheduled into a lab stream in SSO. While lab attendance is not compulsory, it is highly recommended you attend your scheduled lab stream. There is teaching staff to support you through the tasks and answer your course-related questions.


## Weekly quizzes (1% each, 10% towards final grade) {-}

There are eleven (11) weekly quizzes due at the end of each week from week 2 - week 12 (due Friday at 10pm). These are worth 1% each,  where your one (1) quiz that has the lowest mark will be dropped from your final grade. This means overall, your weekly quizzes will total 10% towards your final grade.
These are designed for you to solidify concepts and practice skills taught each week and thereby keep up with the material.
There are practice quizzes for you to attempt prior to sitting the weekly quiz, these are not assessed. There is also a practice quiz for week 1.
\
\


*We are looking forward to meeting you and working with you all this semester.*